---
title: 'Mapping #GoldOpen Tweets'
author: ericayee
date: "September 3, 2018"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I've wanted to learn R for a while for its data analysis capabilities, so here's my first project in R based on this [Storybench tutorial](http://www.storybench.org/build-animated-map-tweets-nba-finals-r/).

When the film "Crazy Rich Asians" came out in theaters on August 15, 2018, at least dozens of celebrities, organizations, and private citizens bought out theaters to ensure a successful "#GoldOpen," as the campaign was called on Twitter. As the first major Hollywood film in 25 years with an all-Asian cast, the film is a cultural moment for Asian AMericans like myself.

I followed this [Storybench tutorial](http://www.storybench.org/build-animated-map-tweets-nba-finals-r/), which has detailed instructions on tracking tweets about the NBA finals, using #GoldOpen tweets instead. The rtweet library captures tweets from up to the last 10 days, so I was too late to see the majority of tweets when the film was released. But in my results I observed tweets from a follow-up campaign for "Searching," which stars John Cho and had a wide release on August 31, 2018.

## Setting up the Twitter app

I followed the instructions on the [rtweet website](https://rtweet.info/) to set up a Twitter app and authenticated it using the `create_token()` function as shown in the documentation.

# Setting up the R environment

Here are the packages I used to pull, analyze, and visualize the tweets.

```{r echo = TRUE, results = 'hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(rtweet)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(ggraph)
library(igraph)
library(hrbrthemes)
library(ggalt)
library(ggthemes)
```

## Let's get some tweets!

I found tweets with the hashtag #GoldOpen using the rtweet package.

```{r echo = TRUE, results = 'hide', message=FALSE, warning=FALSE}
GoldOpen <- rtweet::search_tweets(
  "GoldOpen",
  n = 50000,
  retryonratelimit = TRUE)
```

Here's the data frame of results.

```{r}
#GoldOpen %>% head(10)
GoldOpen %>% dplyr::glimpse()
```


The data frame has 2,000 observations and 88 variables. Let's see which variables from the query are available when I use the `users_data()` function.

```{r}
tibble::as_tibble(
  base::intersect(x = base::names(GoldOpen), 
                                  y = base::names(rtweet::users_data(GoldOpen))))
```

I can store those results when I look at the user data of the #GoldOpen query in a new data frame.

```{r}
UsersGoldOpen <- rtweet::users_data(GoldOpen)
UsersGoldOpen %>% head(10)
```

Now that I've looked at user data, I move on to tweet data.

```{r}
tibble::as_tibble(
  intersect(x = base::names(GoldOpen), 
                                  y = base::names(rtweet::tweets_data(GoldOpen))))
```

Again, I can save the results of running that function on the GoldOpen query to a dataframe.

```{r}
TweetsGoldOpen <- rtweet::tweets_data(GoldOpen)
```

Here are the contents of the first 10 tweets stored in the `text` variable.

```{r}
GoldOpen %>%
    dplyr::select(text) %>%
    utils::head(10)
```

## Construct a timeline of tweets

I can plot tweets over time using the `rtweet::ts_plot()` function and a few other packages to tweak the visuals.

```{r}
gg_ts_plot <- GoldOpen %>%
    rtweet::ts_plot(., by = "15 minutes") +
    ggthemes::theme_gdocs() +
    ggplot2::theme(plot.title =
        ggplot2::element_text(face = "bold")) +
    ggplot2::labs(
        x = NULL,
        y = NULL,
        title = "Freequency of #GoldOpen tweets",
        subtitle = "Tweet counts aggregated using 15-min intervals",
        caption = "\nSource: Data collected from Twitter's REST API via rtweet")
```

Here's the plot.

```{r}
gg_ts_plot
```

## Adding geographic information

Now that I've seen time information, let's add latitude and longitutde to the tweet data.

```{r}
GoldOpenLoc <- rtweet::lat_lng(GoldOpen)
```

I can verify the function added lat and lng variables to the GoldOpen data frame.

```{r}
GoldOpenLoc %>% names() %>% tail(2)
```

Let's see how many tweets have latitude and longitude information.

```{r}
GoldOpenLoc %>% dplyr::distinct(lng) %>% base::nrow()
```

Only 26 of the 2,000 tweets have the geographic information, but we can map the results we have. Let's rename lng to long, which will be helpful later on.

```{r}
GoldOpenLoc <- GoldOpenLoc %>% dplyr::rename(long = lng)
```

## Creating a world map of tweets

I first need to get the data for a world map.

```{r}
World <- ggplot2::map_data("world")
World %>% head(10)
```

This next function will create a map using that data.

```{r}
ggWorldMap <- ggplot2::ggplot() + 
    ggplot2::geom_polygon(data = World,
        aes(x = long,
        y = lat,
        group = group),
        fill = "grey82",
        color = "white",
        alpha = 0.6)

ggWorldMap +
    ggplot2::theme_void() +
    ggplot2::labs(title = "Basic World Map (geom_polygon)")
```

I can layer the tweet data on top of the world map.

```{r}
gg_Merc_title <- "#GoldOpen tweets worldwide, Aug 25 - Sept 3, 2018"
gg_Merc_cap <- "Tweets collected with rtweet the hashtag #GoldOpen"
gg_mercator_goldopen <- ggWorldMap +
    ggplot2::coord_quickmap() +
    ggplot2::geom_point(data = GoldOpenLoc,
                        aes(x = long, y = lat),
                        size = 1,
                        color = "gold") +
    ggplot2::labs(title = gg_Merc_title,
        caption = gg_Merc_cap) +   
    ggplot2::theme_void()
```

Here's the map with the tweet data.

```{r echo = TRUE, results = 'hide', message=FALSE, warning=FALSE}
gg_mercator_goldopen
```

I now want to zoom in on just tweets from the U.S. The code is almost exactly the same as above. Here is the blank map of the U.S.

```{r}
USmap <- ggplot2::map_data("usa")

ggUSMap <- ggplot2::ggplot() +
    ggplot2::geom_polygon(data = USmap,
        aes(x = long,
        y = lat,
        group = group),
        fill = "grey82",
        color = "grey92",
        alpha = 0.6)
        
ggUSMap + ggplot2::theme_void()

```

I can filter the data down to just tweets from the U.S.

```{r}
USfilter <- GoldOpenLoc %>% filter(country == "United States")

USfilter %>% head(10)
```

Then I put these filtered tweets onto the map.

```{r}
gg_USMerc_title <- "#GoldOpen tweets, Aug 25 - Sept 3, 2018"

gg_USMerc_cap <- "Tweets collected with rtweet the hashtag #GoldOpen"

gg_usmercator_goldopen <- ggUSMap +
    ggplot2::coord_quickmap() +
    ggplot2::geom_point(data = USfilter,
        aes(x = long, y = lat),
        size = 1.0,
        color = "gold") +
    ggplot2::labs(title = gg_USMerc_title,
        caption = gg_USMerc_cap) +
    ggplot2::theme_void()
    
gg_usmercator_goldopen

```

There's not much I can extrapolate from such a small sample size, but I do see the few tweets I did get concentrated in urban coastal areas (SF, LA, NYC), where many Asians live.